How topos theory can help commutative algebra

### 2018-07-06 slides-como2018.tex
Notes finished

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\AA}{\mathbb{A}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ull}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ull{\AA}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}{\vdash_{\!\!\!\vec x}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates

@@

      \hil{Background on the internal language}
      \medskip
    \end{column}
    \begin{column}{0.33\textwidth}
      \centering
      \bignumber{2}
      \smallskip

      \hil{Applications in commutative algebra}
      \medskip
    \end{column}
    \begin{column}{0.33\textwidth}
      \centering
      \bignumber{3}
      \smallskip

@@

  Using the internal language of toposes, we will give a short, conceptual and
  constructive proof of Grothendieck's generic freeness lemma. Again, if so
  desired, one can unwind the internal proof to obtain a
  constructive proof which doesn't reference topos theory. The proof obtained
  in this way is still an improvement on the previously known proofs, requiring
  no advanced prerequisites in commutative algebra, and takes
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/paper-generic-freeness.pdf}{about
  a page}.}

@@

  Details on the internal proof are in Section~11.5 of
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{these
  notes}. The
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/paper-generic-freeness.pdf}{external
  proof} obtained by unwinding the internal one is still quite direct, compared
  to the previously published proofs, and interestingly follows a curious
  course: It starts with verifying, in an inductive manner, that~$B$ and~$M$
  are free; that~$A \to B$ is of finite presentation; and that~$M$ is finitely
  presented as a~$B$-module. Then the assumption for~$f = 1$ is used, rendering
  the prior steps moot, since over the zero ring any module is free and any
  algebra is of finite presentation. External translations of internal proofs
  which use double negation will always take such a course.}

@@

@@

  \jnote{1}{\footnotesize At the
  \fixedhref{https://sbseminar.wordpress.com/2009/08/06/algebraic-geometry-without-prime-ideals/}{Secret
  Blogging Seminar}, there was an insightful long-running discussion on the
  merits of the two approaches. Two disadvantages of the approach using
  locally ringed spaces is that the underlying topological spaces don't
  actually parametrize ``honest'', ``geometric'' points, but irreducible closed
  subsets; and that they don't work well in a constructive setting. (For this,
  they would have to be replaced by locally ringed locales.)

  The functorial approach is more economical, philosophically rewarding, and
  works constructively. Given a functor~$F : \Ring \to \Set$, we imagine~$F(A)$
  to be the set of~``$A$-valued points'' of the hypothetical scheme described
  by~$F$, the set of ``points with coordinates in~$A$''. These sets have direct
  geometric meaning. However, typically only field-valued points are easy
  to describe. For instance, the functor representing projective~$n$-space is
  given on fields by
  \[ \begin{array}{r@{}c@{}l}
  K &{}\longmapsto{}& \text{the set of lines through the origin in~$K^{n+1}$} \\
  && \qquad \cong \{ [x_0 \hg \cdots \hg x_n] \,|\, \text{$x_i \neq 0$ for some~$i$} \},
  \end{array} \]
  whereas on general rings it is given by
  \[ \begin{array}{r@{}c@{}l}
    A &{}\longmapsto{}& \text{the set of quotients~$A^{n+1} \twoheadrightarrow P$,
    where~$P$ is projective,} \\
    && \qquad \text{modulo isomorphism}.
  \end{array} \]
  It are these more general kinds of points which impart a meaningful sense of
  cohesion on the field-valued points, so they can't simply be dropped from
  consideration.}

@@

  A \hil{finitely presented synthetic scheme} is a set which can be covered by
  finitely many qc-open f.p. synthetic affine schemes~$U_i$ such that the
  intersections~$U_i \cap U_j$ can be covered by finitely many qc-open f.p.
  synthetic affine schemes.
\end{frame}

@@

  \begin{itemize}
  \justifying
    \item Internally to~$\Zar(S)$, $\Spec(S)$ can be constructed as the
    \hil{largest subtopos} where~$\flat \affl \to \affl$ is bijective.
    \item Internally to~$\Spec(S)$, $\Zar(S)$ can be constructed as the
    \hil{classifying topos} of local~$\O_S$-algebras which are local over~$\O_S$.
    \item $\Zar(A)$ is the \hil{lax pullback}~$(\Set
    \Rightarrow_{\Set[\mathrm{Ring}]} \Set[\mathrm{LocRing}])$.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Properties of the affine line}
  \begin{itemize}
    \item $\affl$ is a local ring:
    \[ 1 \neq 0 \qquad\qquad \text{$x + y$ \inv} \Longrightarrow \text{$x$ \inv} \vee
    \text{$y$ \inv} \]
    \item $\affl$ is a field:
    \begin{align*}
      \neg(x = 0) &\Longleftrightarrow \text{$x$ invertible \quad [Kock 1976]} \\
      \neg(\text{$x$ invertible}) &\Longleftrightarrow \text{$x$ nilpotent}
    \end{align*}
    \item $\affl$ satisfies the axiom of microaffinity: Any map $f : \Delta \to
    \affl$ is of the form~$f(\varepsilon) = a + b \varepsilon$ for unique
    values~$a,b \? \affl$, where~$\Delta = \{ \varepsilon \? \affl \,|\,
    \varepsilon^2 = 0 \}$.
    \item Any map $\affl \to \affl$ is a polynomial.
    \item $\affl$ is anonymously algebraically closed: Any monic polynomial
    does \emph{not not} have a zero.
  \end{itemize}
\end{frame}

@@

% \emph{Example.} Let~$x \? \affl$ such that~$x \neq 0$. Set~$R = \affl/(x)$.
% Then~$\Spec(R) = \emptyset$. Thus~$(\affl)^{\Spec(R)}$ is a singleton.
% Hence~$R = 0$. Therefore~$x$ is invertible.
\end{frame}

\begin{frame}{Classifying toposes in algebraic geometry}
  \small
  \centering
  \begin{tabular}{ll}
    \toprule
    (Big) topos & classified theory \\\midrule
    Zariski & local rings [Hakim 1972] \\[0.6em]
    étale & separably closed local rings [Hakim 1972, Wraith 1979] \\[0.6em]
    \usebeamercolor[fg]{item}{fppf} & \usebeamercolor[fg]{item}{fppf-local rings} \\
    & \textcolor{red!90}{(conjecturally: algebraically closed local rings)} \\[0.6em]
    \textcolor{red!80}{ph} & \textcolor{red!80}{?? (conjecturally: algebraically closed valuation rings} \\
    & \qquad \textcolor{red!80}{validating the projective Nullstellensatz)} \\[0.6em]
    \usebeamercolor[fg]{item}{surjective} & \usebeamercolor[fg]{item}{algebraically closed geometric fields} \\[0.6em]
    \textcolor{red!80}{$\neg\neg$} & \textcolor{red!80}{?? (conjecturally: algebraically closed geometric} \\
    & \qquad \textcolor{red!80}{fields are integral over the base)} \\[0.6em]
    \usebeamercolor[fg]{item}{infinitesimal} & \usebeamercolor[fg]{item}{local algebras together with a nilpotent ideal [Hutzler 2018]} \\[0.6em]
    \textcolor{red!100}{crystalline} & \textcolor{red!100}{??} \\
    \bottomrule
  \end{tabular}
\end{frame}

@@

\begin{frame}{The Kripke--Joyal semantics for the little Zariski topos}
  Recall~$A[f^{-1}] = \bigl\{ \frac{u}{f^n} \,|\, u \in A, n \in \NN \bigr\}$.

  \begin{itemize}
    \item $\Sh(\Spec(A)) \models \text{``For all~$x \in A^\sim$, \ldots''}$

          Meaning: For all~$f \in A$ and all~$x \in A[f^{-1}]$, \ldots
          \medskip

    \item $\Sh(\Spec(A)) \models \text{``There is~$x \in A^\sim$ such that \ldots''}$

          \mbox{Meaning: There is a partition of unity,~$1 = f_1 + \cdots + f_n \in A$,}
          such that for each~$i$, there exists~$x_i \in A[f_i^{-1}]$
          with~\ldots
          \medskip

    \item $\Sh(\Spec(A)) \models \text{``$\varphi$ implies $\psi$''}$

          Meaning: For all~$f \in A$, if~$\varphi$ on stage~$f$, then~$\psi$ on
          stage~$f$.
  \end{itemize}
\end{frame}

>>>>

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\AA}{\mathbb{A}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LocRing}{\mathrm{LocRing}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ull}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ull{\AA}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}{\vdash_{\!\!\!\vec x}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates

@@

      \hil{Background on the internal language}
      \medskip
    \end{column}

    \begin{column}{0.33\textwidth}
      \centering
      \bignumber{2}
      \smallskip

      \hil{Applications in commutative algebra}
      \medskip
    \end{column}

    \begin{column}{0.33\textwidth}
      \centering
      \bignumber{3}
      \smallskip

@@

  Using the internal language of toposes, we will give a short, conceptual and
  constructive proof of Grothendieck's generic freeness lemma. Again, if so
  desired, one can unwind the internal proof to obtain a
  constructive proof which doesn't reference topos theory. The proof obtained
  in this way is still an improvement on the previously known proofs, requiring
  no advanced prerequisites in commutative algebra, and takes
  \fixedhref{https://arxiv.org/abs/1807.01231}{about
  a page}.}

@@

  Details on the internal proof are in Section~11.5 of
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{these
  notes}. The
  \fixedhref{https://arxiv.org/abs/1807.01231}{external
  proof} obtained by unwinding the internal one is still quite direct, compared
  to the previously published proofs, and interestingly follows a curious
  course: It starts with verifying, in an inductive manner, that~$B$ and~$M$
  are free; that~$A \to B$ is of finite presentation; and that~$M$ is finitely
  presented as a~$B$-module. Then the assumption for~$f = 1$ is used, rendering
  the prior steps moot, since over the zero ring any module is free and any
  algebra is of finite presentation. External translations of internal proofs
  which use double negation will always take such a course.}

@@

  \jnote{1}{One doesn't need to be an expert in topos theory in order to know
  that many notions in algebraic geometry are inspired by notions in algebra
  and that proofs in algebraic geometry often proceed by reducing to algebra.
  The internal language is a way of making this connection precise: In many
  cases, the former are simply interpretations of the latter internal
  to~$\Sh(X)$. Because this connection is precise instead of informal,
  additional value is gained: We can skip many basic proofs in algebraic
  geometry because they're just externalizations of proofs in algebra carried
  out internally to~$\Sh(X)$.}

  \jnote{2}{A basic example is as follows. A short exact sequence of sheaves of
  modules looks like a short exact sequence of plain modules from the internal
  point of view of~$\Sh(X)$. If the two outer sheaves are of finite type, then
  from the internal point of view, the two outer modules will look like
  finitely generated modules. Because the standard proof of the proposition
  quoted on the lower right is intuitionistically valid, it follows that, from
  the internal point of view, the middle module is too finitely generated.
  Consulting the dictionary a second time, this amounts to saying that the
  middle sheaf is of finite type.

  More details on this research program can be found in
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{these
  notes}, partly reported on at the
  \fixedhref{https://www.youtube.com/watch?v=7S8--bIKaWQ}{2015 IHÉS
  conference}. Even though many important dictionary entries
  are still missing (for instance pertaining derived categories and
  intersection theory), I believe that it is already at its current stage
  useful to working algebraic geometers.

  The next few slides illustrate an entirely different way of approaching
  algebraic geometry using topos theory.}

@@

  \jnote{1}{\footnotesize At the
  \fixedhref{https://sbseminar.wordpress.com/2009/08/06/algebraic-geometry-without-prime-ideals/}{Secret
  Blogging Seminar}, there was an insightful long-running discussion on the
  merits of the two approaches. Two disadvantages of the approach using
  locally ringed spaces is that the underlying topological spaces don't
  actually parametrize ``honest'', ``geometric'' points, but the more complex
  notion of irreducible closed subsets; and that they don't work well in a
  constructive setting. (For this, they would have to be replaced by locally
  ringed locales.)

  The functorial approach is more economical, philosophically rewarding, and
  works constructively. Given a functor~$F : \Ring \to \Set$, we imagine~$F(A)$
  to be the set of~``$A$-valued points'' of the hypothetical scheme described
  by~$F$, the set of ``points with coordinates in~$A$''. These sets have direct
  geometric meaning. However, typically only field-valued points are easy
  to describe. For instance, the functor representing projective~$n$-space is
  given on fields by
  \[ \begin{array}{r@{}c@{}l}
  K &{}\longmapsto{}& \text{the set of lines through the origin in~$K^{n+1}$} \\
  && \qquad \cong \{ [x_0 \hg \cdots \hg x_n] \,|\, \text{$x_i \neq 0$ for some~$i$} \},
  \end{array} \]
  whereas on general rings it is given by
  \[ \begin{array}{r@{}c@{}l}
    A &{}\longmapsto{}& \text{the set of quotients~$A^{n+1} \twoheadrightarrow P$,
    where~$P$ is projective,} \\
    && \qquad \text{modulo isomorphism}.
  \end{array} \]
  It is these more general kinds of points which impart a meaningful sense of
  cohesion on the field-valued points, so they can't simply be dropped from
  consideration.}

@@

  A \hil{finitely presented synthetic scheme} is a set which can be covered by
  finitely many qc-open f.p. synthetic affine schemes~$U_i$ such that the
  intersections~$U_i \cap U_j$ can be covered by finitely many qc-open f.p.
  synthetic affine schemes.

  \jnote{1}{In the internal universe of the big Zariski topos of a base
  scheme~$S$, $S$-schemes can simply be modeled by sets (enjoying the special
  property that, in a certain precise sense, they are locally affine). This
  slide expresses some of the basic constructions of~$S$-schemes in that
  language.

  Particularly nice are the following items.
  \begin{itemize}\justifying
    \item Projective~$n$-space can be given by the any of the two quite naive
    expressions displayed on the slide.
    \item Let~$X$ be an~$S$-scheme. We often think about a sheaves
    of~$\O_X$-modules over~$X$ by their fibers; but for a rigorous
    treatment in the standard foundations, we have to take the full sheaf
    structure into account; the fibers do not determine a sheaf
    uniquely.

    From the internal point of view of~$\Zar(S)$, a sheaf of~$\O_X$-modules is
    indeed simply a family of~$\affl$-modules, one~$\affl$-module for each
    element of~$\ull{X}$. The slide illustrates how we can define the Serre
    twisting sheaves in this language.
  \end{itemize}}

  \jnote{2}{\begin{itemize}\justifying
    \item The spectrum of an~$\affl$-algebra can be given by the naive
    expression displayed on the slide. It looks like this expression can't be
    right, ignoring any non-maximal ideals; however, it is.
    \item The big Zariski topos of an~$S$-scheme~$X$ is, from the internal
    point of view of~$\Zar(S)$, simply the slice topos~$\Set/\ull{X}$. Hence to
    give an~$X$-scheme simply amounts to giving an~$\ull{X}$-indexed family of
    sets.
  \end{itemize}

  Synthetic algebraic geometry has been developed up to the point of étale
  geometric morphisms. Much remains to be done: For instance, as of yet there
  is only an account of Čech methods for computing cohomology, there is not yet
  a synthetic treatment of true cohomology. Derived categories and intersection
  theory are also missing.}
\end{frame}

@@

  \begin{itemize}
  \justifying
    \item Internally to~$\Zar(S)$, $\Spec(S)$ can be constructed as the
    \hil{largest subtopos} where~$\flat \affl \to \affl$ is bijective.
    \item Internally to~$\Spec(S)$, $\Zar(S)$ can be constructed as the
    \hil{classifying topos} of local~$\O_S$-algebras which are local over~$\O_S$.
    \item $\Zar(A)$ is the \hil{lax pullback}~$(\Set
    \Rightarrow_{\Set[\Ring]} \Set[\LocRing])$.
  \end{itemize}

  \jnote{1}{Let~$A$ be a ring. By definition, we obtain a geometric
  morphism~$\Set \to \Set[\Ring]$ into the classifying topos of rings.
  There is also a geometric morphism~$\Set[\LocRing] \to \Set[\Ring]$, obtained
  by realizing that any local ring is in particular a ring. These morphisms fit
  together in a lax pullback square as follows:
  \[ \xymatrix{
    \Zar(A) \ar[r] \ar[d] & \Set[\LocRing] \ar[d] \\
    \Set \ar[r] \ar@{}[ur]^(.3){}="a"^(.7){}="b" \ar@{=>} "a";"b" & \Set[\Ring]
  } \]
  This observation is joint with Peter Arndt and Matthias Hutzler.

  Incidentally, the pseudo pullback of the morphism~$\Set[\LocRing] \to
  \Set[\Ring]$ along~$\Set \to \Set[\Ring]$ is not very interesting: It's the
  largest subtopos of~$\Set$ where~$A$ is a local ring. Assuming the law of
  excluded middle, this subtopos is either the trivial topos (if~$A$ is not
  local) or~$\Set$ (if~$A$ is local).

  There is also a way of realizing the little Zariski topos of~$A$ as a pseudo
  pullback, exploiting that the (localic) spectrum construction is geometric.
  See Section~12.6 of
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{these
  notes} for details.
  }
\end{frame}

@@

\begin{frame}{Properties of the affine line}
  \begin{itemize}
    \item $\affl$ is a local ring:
    \[ 1 \neq 0 \qquad\qquad \text{$x + y$ \inv} \Longrightarrow \text{$x$ \inv} \vee
    \text{$y$ \inv} \]
    \item $\affl$ is a field:
    \begin{align*}
      \neg(x = 0) &\Longleftrightarrow \text{$x$ invertible \quad [Kock 1976]} \\
      \neg(\text{$x$ invertible}) &\Longleftrightarrow \text{$x$ nilpotent}
    \end{align*}
    \item $\affl$ satisfies the axiom of microaffinity: Any map $f : \Delta \to
    \affl$ is of the form~$f(\varepsilon) = a + b \varepsilon$ for unique
    values~$a,b \? \affl$, where~$\Delta = \{ \varepsilon \? \affl \,|\,
    \varepsilon^2 = 0 \}$.
    \item Any map $\affl \to \affl$ is a polynomial.
    \item $\affl$ is anonymously algebraically closed: Any monic polynomial
    does \emph{not not} have a zero.
  \end{itemize}

  \jnote{1}{The axiom of microaffinity is a special instance of the
  \emph{Kock--Lawvere axiom} known from synthetic differential geometry. We'll
  see on the next slide that~$\affl$ validates an unusually strong form of the
  Kock--Lawvere axiom, not at all satisfied in the usual well-adapted models of
  synthetic differential geometry.

  The fact that, internally to~$\Zar(S)$, any map~$\affl \to \affl$ is a
  polynomial can be seen as a formal version of the general motto that in
  algebraic geometry, ``morphisms are polynomials''.}
\end{frame}

@@

  \jnote{1}{Let~$R$ be an~$\affl$-algebra. An element~$f \in R$ induces
  an~$\affl$-valued function on~$\Spec(R)$; functions of this form can
  reasonably be called ``algebraic''. In a synthetic context, there should be
  no other~$\affl$-valued functions on~$\Spec(R)$ as these algebraic ones, and
  different algebraic expressions should yield different functions. This is
  precisely what the bijectivity of the displayed map expresses (in a positive
  way).

  In synthetic differential geometry, the closest cousin of synthetic algebraic
  geometry, the analogue of the displayed map is only bijective for Weil
  algebras such as~$\affl[X]/(X^2)$ or $\affl[X,Y]/(X^2,XY)$, not for arbitrary
  finitely presented~$\affl$-algebras. This is a major difference to synthetic
  differential geometry.}

  \jnote{2}{The notion of synthetic quasicoherence is interesting for a number
  of reasons:\vspace*{-1.2em}
  \begin{itemize}\justifying
    \item All currently known properties of~$\affl$, such as all the properties
    listed on the previous slide, follow from the statement that~$\affl$ is
    synthetically quasicoherent.

    For instance, here is how we can verify the field property. Let~$x \?
    \affl$ such that~$x \neq 0$. Set~$R = \affl/(x)$. Then~$\Spec(R) =
    \emptyset$. Thus~$(\affl)^{\Spec(R)}$ is a singleton. Hence~$R = 0$.
    Therefore~$x$ is invertible.
    \item Given an~$\affl$-module~$E$, we can formulate the following variant
    of the axiom of synthetic quasicoherence: ``For any finitely
    presented~$\affl$-algebra~$R$, the canonical map~$R \otimes_\affl E \to
    E^{\Spec(R)}$ is bijective.'' This axiom is satisfied if and only if~$E$ is
    induced by a quasicoherent sheaf of~$\O_S$-modules.
    \item The notion of synthetic quasicoherence is central to synthetic
    algebraic geometry. The notions of synthetic open immersions, closed
    immersion, schemes and several others all refer to synthetic
    quasicoherence.
  \end{itemize}}

  \jnote{3}{An anlogue of synthetic quasicoherence holds in the classifying
  topos of rings, demonstrating that even presheaf toposes can validate
  interesting nontrivial nongeometric sequents.

  We believe that an analogue of synthetic quasicoherence holds for the generic
  model of any geometric theory. This is work in progress. If true, this would
  yield a major source of nongeometric sequents in classifying toposes. Because
  of the many applications on nongeometric sequents, it's very desirable to
  possess such a source.

  The mystery of nongeometric sequents is this: On the one hand, they are very
  useful to have because of surprising applications; on the other hand, they
  are as of yet quite elusive.}
\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses-lighter}\end{minipage}}
\begin{frame}{Classifying toposes in algebraic geometry}
  \small
  \centering
  \begin{tabular}{ll}
    \toprule
    (big) topos & classified theory \\\midrule
    Zariski & local rings [Hakim 1972] \\[0.6em]
    étale & separably closed local rings [Hakim 1972, Wraith 1979] \\[0.6em]
    \usebeamercolor[fg]{item}{fppf} & \usebeamercolor[fg]{item}{fppf-local rings} \\
    & \textcolor{red!90}{(conjecturally: algebraically closed local rings)} \\[0.6em]
    \textcolor{red!80}{ph} & \textcolor{red!80}{?? (conjecturally: algebraically closed valuation rings} \\
    & \qquad \textcolor{red!80}{validating the projective Nullstellensatz)} \\[0.6em]
    \usebeamercolor[fg]{item}{surjective} & \usebeamercolor[fg]{item}{algebraically closed geometric fields} \\[0.6em]
    \textcolor{red!80}{$\neg\neg$} & \textcolor{red!80}{?? (conjecturally: algebraically closed geometric} \\
    & \qquad \textcolor{red!80}{fields which are integral over the base)} \\[0.6em]
    \usebeamercolor[fg]{item}{infinitesimal} & \usebeamercolor[fg]{item}{local algebras together with a nilpotent ideal [Hutzler 2018]} \\[0.6em]
    \textcolor{red!100}{crystalline} & \textcolor{red!100}{??} \\
    \bottomrule
  \end{tabular}

  \jnote{1}{Toposes, and also more specifically classifying toposes, originated
  in algebraic geometry. It is therefore deeply embarassing that as of now,
  still very little is known about the theories classified by the major toposes
  in active use by algebraic geometers.
  \begin{center}
    \includegraphics[width=0.99\textwidth]{wraith-generic-galois-theory} \\
    \footnotesize
    Gavin Wraith. Generic Galois theory of local rings. 1979.
  \end{center}}

  \jnote{2}{For almost forty years, only the big Zariski topos and its étale
  subtopos were understood in that way.
  \fixedhref{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{These~2017
  notes} answer the question for the fppf topology and the surjective topology
  (in Section~21) and state conjectures for the ph topology and the
  double negation topology. However, while good to have, the answer for the
  fppf topology remains unsatisfactory, since Wraith's conjecture that the fppf
  topos classifies the simpler theory of algebraically closed local rings has
  neither been confirmed nor refuted.

  A couple of weeks ago, Matthias Hutzler managed to determine the theory
  classified by the big infinitesimal topos of a ring~$A$: It classifies
  pairs~$(B,\aaa)$ consisting of a local~$A$-algebra~$B$ and a nilpotent
  ideal~$\aaa \subseteq B$. Details will be in his forthcoming Master's thesis.
  He is currently working on answering the question for the closely related big
  crystalline topos.

  It will be exciting to learn what the crystalline topos and the many other
  toposes in algebraic geometry classify; how algebraic geometry can profit
  from these discoveries; and which new flavors of synthetic algebraic geometry
  they unlock.}
\end{frame}}

@@
